---
title: "hourly_ops_charts.Rmd"
author: "team alpha"
date: "5/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## prep
### import libs
```{r}
library(here)
library(tidyverse)
library(readxl)
library(viridis)
library(lubridate)
library(glue)
```

### set up colors
```{r}
eggplant <- "#403891FF"

my_colors <- function (palette = "dis"){
    dis.palette <- c("#a6cee3", "#ff7f00")
    rcb.palette <- rev(dis.palette)
    two.palette <- c("#EB8055FF", # green #EB8055FF", "#403891FF","#E8FA5BFF"
                     "#403891FF") # eggplant
    if (palette == "dis") 
        return(dis.palette)
    else if (palette == "rcb") 
        return(rcb.palette)
    else if (palette == "two") 
        return(two.palette)
    else stop("Choose cb, rcb, or two only.")
}
```

### Set up fonts
```{r}
library(showtext)
## Loading Google fonts (http://www.google.com/fonts)
font_add_google("spectral", "spectral")
font_add_google("roboto slab", "roboto")

## Automatically use showtext to render text for future devices
showtext_auto()

theme_set(theme_minimal() + theme(text=element_text(size=16,family="roboto"), 
              plot.title = element_text(size=22), 
              plot.margin = unit(c(1,1,1,1), "cm")))
```


## Arrival & delay density plots
### Import data
```{r}
samp <- read.csv("/Users/james/Downloads/759366606_T_ONTIME_REPORTING.csv")
```

### Munge & plot
```{r}
samp2 <- samp %>%
  dplyr::select(ARR_DELAY, DEP_DELAY) %>%
  gather(delay_type, minutes) %>%
  mutate(delay_type = recode(delay_type, "ARR_DELAY" = "Arrival delays", 
                             "DEP_DELAY" = "Departure delays"))

ggplot(samp2, aes(x = delay_type, y = minutes, fill = delay_type, color = delay_type)) +
  scale_fill_manual(values = my_colors("two")) +
  scale_color_manual(values = my_colors("two")) +
  scale_y_continuous(limits = c(-50, 120), breaks = seq(-50, 130, 50)) +
  scale_x_discrete(labels = scales::wrap_format(5)) +
  geom_violin() +
  coord_flip() +

  labs(title = "", 
       x="", y = "Length of delay in minutes") +
  guides(color = FALSE, fill = FALSE)
```

## Hourly dotplots
### One year at a time
```{r fig.height=4, fig.width=4}

plot_hourly_1y <- function(file_name, year_idx, title, x_max, y_max){
  path <- here::here("data", "hourly", file_name)
  atl <- read.csv(path)
  names(atl) <- c("year", "month", "day", "hour", "arrs", "arrs_del", "deps", "deps_del")
  
  atl <- atl %>%
  mutate(prop_del = ( (arrs_del+deps_del) / (arrs+deps)), 
         arrs_on_time = arrs - arrs_del, 
         deps_on_time = deps - deps_del) %>%
  select(-arrs_del, -deps_del) %>%
  filter(hour > 6, hour < 12, 
         year == year_idx) %>%
  mutate(delay_cat = ifelse(prop_del < .1, "90% on time", 
                           ifelse(prop_del < .2, "80% on time", "Worse than 80%"))) 
  
  ggplot(atl, aes(x = deps, y = arrs)) +
  #theme_minimal() +
  scale_x_continuous(limits = c(0, x_max)) +
  scale_y_continuous(limits = c(0, y_max)) +
    scale_color_manual("", 
                breaks = c("90% on time", "80% on time",
                           "Worse than 80%"),
                values = c( "#ff6699",
                            "#403891FF", 
                            
                             
                            "#cccc00"
                           ), 
                labels = c("90% on time", "80% on time",
                           "Worse than 80%")) +
    #scale_color_manual(values = c("#EB8055FF", "#403891FF","#E8FA5BFF")) +
  geom_jitter(aes(color = delay_cat), alpha = 1, size = 1, width = 0.5, height = 0.5) +
  #geom_bin2d(bins = 40) +
    labs(title = glue(title), 
       color = "Hourly performance", 
       x = "Arrivals", 
       y = "Departures") +
    guides(color = FALSE)
  
}

#plot_hourly_1y("atl.csv", 2004, "Hourly Operations at ATL")
plot_hourly_1y("atl.csv", 2005, "Pre-Expansion ATL: 2005", 120, 120)
plot_hourly_1y("atl.csv", 2008, "Post-Expansion ATL: 2008", 120, 120)

plot_hourly_1y("stl.csv", 2005, "Pre-Expansion STL: 2005", 25, 25)
plot_hourly_1y("stl.csv", 2008, "Post-Expansion STL: 2008", 25, 25)
#plot_hourly_1y("atl.csv", 2016, "Hourly Operations at ATL")
#plot_hourly_1y("atl.csv", 2018, "Hourly Operations at ATL")
#plot_hourly_1y("atl.csv", 2014, "Hourly Operations at ATL")

?ggsave

ggsave("../plots/atl_hourly_2004.jpeg", width = 4, height = 4, 
       device = "jpeg", dpi = 300)
```

### Faceted by year
```{r fig.height=4, fig.width=6}
plot_hourly <- function(file_name, title){
  path <- here::here("data", "hourly", file_name)
  atl <- read.csv(path)
  names(atl) <- c("year", "month", "day", "hour", "arrs", "arrs_del", "deps", "deps_del")
  
  atl <- atl %>%
  mutate(prop_del = ( (arrs_del+deps_del) / (arrs+deps)), 
         arrs_on_time = arrs - arrs_del, 
         deps_on_time = deps - deps_del) %>%
  select(-arrs_del, -deps_del)
  
  plot_atl <- atl %>%
  filter(hour > 6, hour < 12) %>%
  filter(year > 2000) %>%
  mutate(delay_cat = ifelse(prop_del < .1, "90% on time", 
                           ifelse(prop_del < .2, "80% on time", "Worse than 80%"))) 
  
  ggplot(plot_atl, aes(x = deps, y = arrs)) +
  scale_x_continuous(limits = c(0, 120), breaks = c(60, 120)) +
    scale_color_manual("", 
                breaks = c("90% on time", "80% on time",
                           "Worse than 80%"),
                values = c( "#ff6699",
                            "#403891FF", 
                            
                             
                            "#cccc00"
                           ), 
                labels = c("90% on time", "80% on time",
                           "Worse than 80%")) +
    geom_point(aes(color = delay_cat), alpha = 1, size = .2) +
  #facet_wrap(~year)  +
  facet_wrap(~year, ncol = 6) +
  labs(title = "", 
       color = "Hourly performance", 
       x = "Arrivals", 
       y = "Departures",
       color = "Hourly on-time-ness") +
    guides(color = FALSE)
  
}

?facet_wrap

#plot_hourly("atl_no_wx.csv", "Hourly Operations Over Time at ATL (bad weather excluded)")
#plot_hourly("atl.csv", "Hourly Operations Over Time at ATL")
#plot_hourly("fll.csv", "Hourly Operations Over Time at FLL")
plot_hourly("atl.csv", "Hourly Operations Over Time at ATL")
```



### ATL: on_timearrivals only
```{r fig.height=5, fig.width=5}
plot_atl <- atl %>%
  filter(hour > 6, hour < 12) 

ggplot(plot_atl, aes(x = deps_on_time, y = arrs_on_time)) +
  scale_x_continuous() +
  geom_point(alpha = .33, size = .33) +
  facet_wrap(~year, nrow = 5, ncol = 3) +
  labs(title = "ATL: On time arrivals and departures only")

ggplot(plot_atl, aes(x = deps, y = arrs)) +
  scale_x_continuous() +
  geom_point(alpha = .33, size = .33) +
  facet_wrap(~year, nrow = 5, ncol = 3) +
  labs(title = "ATL: all arrivals and departures")
```


### ATL proportion chart
```{r}
plot_atl_2 <- plot_atl %>%
  group_by(year, delay_cat) %>%
  summarize(n = n() ) %>%
  mutate(prop = n/sum(n))

ggplot(plot_atl_2, aes(x = year, y = prop , group = delay_cat, color = delay_cat)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  geom_line() +
  labs(title = "Proportion of Timely Hours at ATL", 
       color = "On-timeness of hours")
```

### ATL proportion chart
```{r}
plot_atl_2 <- plot_atl %>%
  group_by(year, delay_cat) %>%
  summarize(n = n() ) %>%
  mutate(prop = n/sum(n))

ggplot(plot_atl_2, aes(x = year, y = prop , group = delay_cat, color = delay_cat)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 1)) +
  geom_line() +
  labs(title = "Proportion of Timely Hours at ATL", 
       color = "On-timeness of hours")
```



### Just delays
```{r}
library(ggridges)
head(ord)

ggplot(ord, aes(x = prop_del, y = as.factor(year))) + 
  geom_density_ridges() +
    labs(title = "Performance by Hour at ORD", 
         x = "Distribution of the Proportion of flights delayed", 
         y = "year")

ggplot(plot_ord, aes(x = year, y = total_delays)) +
    geom_bar(stat = "identity")

ggplot(plot_ord, aes(x = deps_del, y = arrs_del)) +
  scale_x_continuous() +
 scale_color_viridis(option = "plasma") +
  geom_point(aes(color = prop_del), alpha = 1, size = .33) +
  facet_wrap(~year, nrow = 5, ncol = 3) +
  labs(title = "Arrivals and Departures at ORD")
```



```{r}
head(atl)

atl_max <- atl %>%
  filter(hour > 6, hour < 12) %>%
  mutate(date = mdy(glue("{month}-", "{day}-", "{year}")),
         ops_on_time = arrs_on_time + deps_on_time) %>%
  group_by(date, year, month) %>%
  mutate(max_ops_per_day = max(ops_on_time)) %>%
  group_by(year) %>%
  mutate(mean_max = mean(max_ops_per_day))
  
  
ggplot(atl_max, aes(x = year, y = mean_max)) +
  geom_


```

## LGA
### import LGA and munge data
```{r}
path <- here::here("data", "hourly", "atl.csv")
atl <- read.csv(path)
names(atl) <- c("year", "month", "day", "hour", "arrs", "arrs_del", "deps", "deps_del")

head(atl)

tes <- atl %>%
  filter(year == 2004)


lga <- lga %>%
  mutate(prop_del = ( (arrs_del+deps_del) / (arrs+deps)), 
         arrs_on_time = arrs - arrs_del, 
         deps_on_time = deps - deps_del, 
         total_delays = deps_del + arrs_del) 

head(lga)
```


## Discontinuity examples
### No discontinuity
```{r}
# Discontinuity Regression with signal:
set.seed(223)
x <- seq(1,20,.25)
y_before <- rnorm(37, -1.5, 1.5) +x[1:37]*(1/2)
y_after <- rnorm(40, -3, 3) + (x[38:77])*4 - 30

example_df <- data.frame(x,c(y_before, y_after))
names(example_df)[2] <- "y"

# Plot of signal:
ggplot(example_df,aes(x=x,y=y))+
  geom_point() +
  geom_point(data=subset(example_df, x <= 10), color = "grey25") +
  geom_smooth(data=subset(example_df, x <= 10), 
              aes(color = "1"), method='lm',formula=y~x,se=T, 
              fill = "grey55") +
  geom_point(data=subset(example_df, x > 10), color = "grey25") +
  geom_smooth(data=subset(example_df, x > 10), 
              aes(color  = "2"), method='lm',formula=y~x,se=T, 
              fill = "grey55") +
  geom_vline(xintercept = 10, linetype='dashed') +
  ggtitle(label = "") +
  scale_color_manual(values = c("1" = my_colors("two")[1],"2" = my_colors("two")[2]),
                     labels = c("Before", "After")) +
  labs(x = "Time", y = "Response", color = "Treatment Period") +
  theme(legend.position="bottom")

before <- lm(y~x, data = example_df, subset = (x <= 10))
summary(before)

after <- lm(y~x, data = example_df, subset = (x > 10))
summary(after)

ggsave("../plots/discont.jpeg", device = "jpeg", width = 6, height = 4, dpi = 300)
```

### No discontinuity
```{r}
# Discontinuity Regression with NO signal:
set.seed(223)
x <- seq(1,20,.25)
y_before <- rnorm(37, -.5, .5) +x[1:37]*(1/3)
y_after <- rnorm(40, -.5, .5) + x[38:77]*(1/3)

example_df <- data.frame(x,c(y_before, y_after))
names(example_df)[2] <- "y"

# Plot of NO signal:
ggplot(example_df,aes(x=x,y=y))+
  geom_point() +
  geom_point(data=subset(example_df, x <= 10), color = "grey25") +
  geom_smooth(data=subset(example_df, x <= 10), 
              aes(color = "1"), method='lm',formula=y~x,se=T, 
              fill = "grey55") +
  geom_point(data=subset(example_df, x > 10), color = "grey25") +
  geom_smooth(data=subset(example_df, x > 10), 
              aes(color  = "2"), method='lm',formula=y~x,se=T, 
              fill = "grey55") +
  geom_vline(xintercept = 10, linetype='dashed') +
  ggtitle(label = "") +
  scale_color_manual(values = c("1" = my_colors("two")[1],"2" = my_colors("two")[2]),
                     labels = c("Before", "After")) +
  labs(x = "Time", y = "Response", color = "Treatment Period") +
  theme(legend.position="bottom")

before <- lm(y~x, data = example_df, subset = (x <= 10))
summary(before)

after <- lm(y~x, data = example_df, subset = (x > 10))
summary(after)

ggsave("../plots/no_discont.jpeg", device = "jpeg", width = 6, height = 4, dpi = 300)
```

## Discontinuity cases
### Load in StandardReport
```{r}
path <- here::here("data", "StandardReport.csv")
standard <- read_csv(path)

# Removes stupid government tallies
standard <- standard %>%
  filter(!str_detect(Date, 'Sub-Total'))

# Converts Month-Year (JUN-13) to Date Format (2013-06-01)
standard$Date <- as.Date(sub("\\-", "01", standard$Date), "%b%d%y")
```
### TPA's "typical curve"
```{r}
# Getting general arrival demand shape by filtering to TPA:
standard_tpa <- standard %>%
  filter(Facility == 'TPA')

# Total On-Time Arrivals (OPS) for TPA:
ggplot(standard_tpa,aes(x=Date,y=On_Time_Arrivals)) +
  scale_y_continuous(labels = scales::comma) +
  geom_point() +
  geom_smooth(color = eggplant) +
  ggtitle(label = "") +
  labs(y = "Total On-Time Flight Arrivals") 

ggsave("../plots/typical_curve.jpeg", device = "jpeg", width = 6, height = 4, dpi = 300)
```

```{r}
  standard_air <- standard %>%
    filter(Facility == "ATL")

  ggplot(standard_air,aes(x=Date,y=Perc_On_Time_Gate_Arrs))+
    geom_point() +
    geom_point(data=subset(standard_sea, Date >= "2009-12-01")) +
    geom_smooth(data=subset(standard_sea, Date >= "2009-12-01"), 
                aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2]) +
    geom_point(data=subset(standard_sea, Date <= "2008-12-01")) +
    geom_smooth(data=subset(standard_sea, Date <= "2008-12-01"), 
                aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1]) +
    geom_vline(xintercept = as.numeric(as.Date("2009-12-01")), linetype='dashed') +
    geom_vline(xintercept = as.numeric(as.Date("2008-12-01")), linetype='dashed') +
    ggtitle(label = "") +
    scale_y_continuous(labels = function(x) paste0(x, "%")) +
    scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                  "2" = my_colors("dis")[1]),
                       labels = c("Before","After")) 
    labs(y = "Percent On-Time", color = "Expansion Period", x = "") +
    annotate("label", x = as.Date("2008-12-01"), y = 66.25, 
             label = "11/20/08", size = 4) +
    theme(axis.text.x = element_text(size = 18), 
          axis.text.y = element_text(size=18)) 
```

### SEA discont
#### on-time arrs
```{r fig.height=4.5, fig.width=6}
standard_air <- standard %>%
  filter(Facility == "FLL")

ggplot(standard_a, aes(x=Date,y=Perc_On_Time_Gate_Arrs))+
  geom_point() +
  geom_point(data=subset(standard_air, Date >= "2009-12-01")) +
  geom_smooth(data=subset(standard_air, Date >= "2009-12-01"), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= "2008-12-01")) +
  geom_smooth(data=subset(standard_air, Date <= "2008-12-01"), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
  geom_vline(xintercept = as.numeric(as.Date("2009-12-01")), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date("2008-12-01")), linetype='dashed') +
  ggtitle(label = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "Percent On-Time", color = "Expansion Period", x = "") +
  annotate("label", x = as.Date("2008-12-01"), y = 66.25, 
           label = "11/20/08", size = 4) +
  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 

ggsave("../plots/fll_arrs.jpeg", width = 4.5, height = 6, 
       device = "jpeg", dpi = 300)
```

#### % arrivals on-time
```{r fig.height=4.5, fig.width=6}

standard_air <- standard %>%
  filter(Facility == "FLL")

before_date <- "2014-08-01"
after_date <- "2015-07-01"

ggplot(standard_air,aes(x=Date,y=On_Time_Arrivals))+
    geom_point() +
  scale_y_continuous(labels = scales::comma_format(scale = .001, suffix = "k") ) +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
  geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
  ggtitle(label = "") +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "On-Time Flight Arrivals", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 


```

### FLL discont
#### on-time arrs
```{r fig.height=4.5, fig.width=6}
  
standard_air <- standard %>%
  filter(Facility == "FLL")

before_date <- "2014-08-01"
after_date <- "2015-07-01"

ggplot(standard_a, aes(x=Date,y=Perc_On_Time_Gate_Arrs))+
  geom_point() +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
    ggtitle(label = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "Percent On-Time", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 

ggsave("../plots/fll_arrs.jpeg", width = 4.5, height = 6, 
       device = "jpeg", dpi = 300)
```

#### % arrivals on-time
```{r}

```

### ATL discont
#### # arrivals on-time
```{r fig.height=4.5, fig.width=6}

standard_air <- standard %>%
  filter(Facility == "ATL")

before_date <- "2006-04-01"
after_date <- "2007-04-01"

ggplot(standard_air,aes(x=Date,y=On_Time_Arrivals))+
    geom_point() +
  scale_y_continuous(labels = scales::comma_format(scale = .001, suffix = "k") ) +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
  geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
  ggtitle(label = "") +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "On-Time Flight Arrivals", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 


```
#### % arrivals on-time
```{r fig.height=4.5, fig.width=6}
standard_air <- standard %>%
  filter(Facility == "ATL")

before_date <- "2006-04-01"
after_date <- "2007-04-01"

ggplot(standard_a, aes(x=Date,y=Perc_On_Time_Gate_Arrs))+
  geom_point() +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
    ggtitle(label = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "Percent On-Time", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 
```

### STL

### ATL discont
#### # arrivals on-time
```{r fig.height=4.5, fig.width=6}

standard_air <- standard %>%
  filter(Facility == 'STL') %>%
  filter(Date >= "2003-11-01")

before_date <- "2006-05-01"
after_date <- "2007-05-01"

ggplot(standard_air,aes(x=Date,y=On_Time_Arrivals))+
    geom_point() +
  scale_y_continuous(labels = scales::comma_format(scale = .001, suffix = "k") ) +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
  geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
  ggtitle(label = "") +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "On-Time Flight Arrivals", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 
```

#### % arrivals on-time
```{r fig.height=4.5, fig.width=6}
standard_air <- standard %>%
  filter(Facility == "STL")

before_date <- "2006-04-01"
after_date <- "2007-04-01"

ggplot(standard_a, aes(x=Date,y=Perc_On_Time_Gate_Arrs))+
  geom_point() +
  geom_point(data=subset(standard_air, Date >= after_date)) +
  geom_smooth(data=subset(standard_air, Date >= after_date), 
              aes(color = "2"), method='lm',formula=y~x,se=F, color = my_colors("dis")[2], 
              size = 2) +
  geom_point(data=subset(standard_air, Date <= before_date)) +
  geom_smooth(data=subset(standard_air, Date <= before_date), 
              aes(color  = "0"), method='lm',formula=y~x,se=F, color = my_colors("dis")[1], 
              size = 2) +
geom_vline(xintercept = as.numeric(as.Date(after_date)), linetype='dashed') +
  geom_vline(xintercept = as.numeric(as.Date(before_date)), linetype='dashed') +
    ggtitle(label = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_color_manual(values = c("0" = my_colors("dis")[2], "1" = "black", 
                                "2" = my_colors("dis")[1]),
                     labels = c("Before","After")) +
  labs(y = "Percent On-Time", color = "Expansion Period", x = "") +

  theme(axis.text.x = element_text(size = 18), 
        axis.text.y = element_text(size=18)) 
```
